# Pre-Class Assignment
# Grace A. O'Brien
### Getting Started

We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 
# Installing gapminder package
```{r}
install.packages("gapminder")
```


# Loading gapminder into R
```{r}
library(dplyr)
library(gapminder)
gapminder
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?
```{r}
names(gapminder)
colSums(table(gapminder$country, gapminder$continent))
c <- gapminder %>%
  group_by(continent) %>%
  distinct(country) %>%
  arrange(continent)
colSums(table(c))
```
ANSWER: Africa- 52, Americas- 25, Asia-33, Europe- 30, Oceania-2.

2. Which European nation had the lowest GDP per capita in 1997? 
```{r}
gapminder %>%
  filter(year == 1997, continent == "Europe") %>%
  arrange(gdpPercap, country)

```
ANSWER: Albania with a gdp per cap of 3,193. 

3. According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
gapminder %>%
  group_by(continent) %>%
  filter(year <= "1989", year >= "1980") %>%
  summarise(avg_lifeExp = mean(lifeExp))
```
ANSWER: Africa-52.5, Americas- 67.2, Asia- 63.7, Europe- 73.2, Oceania- 74.8.

4. What 5 countries have the highest total GDP over all years combined?
```{r}
gapminder %>%
  group_by(country) %>%
  summarise(total_gdp = sum(gdpPercap)) %>%
  arrange(desc(total_gdp))
```
ANSWER: Kuwait, Switzterland, Norway, United States, Canada. 

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._
```{r}
x <- gapminder %>%
  filter(lifeExp >= "80") %>%
  select(country, lifeExp, year)
x
```
ANSWER: There are 22 country/year combinations in which life expectancy was at least 80 years- Australia(2002, 2007), Canada(2007), France(2007), Hong Kong, China (1997, 2002, 2007), Iceland(2002, 2007), Israel (2007), Italy (2002, 2007), Japan (1997, 2007), New Zealand (2007), Norway (2007), Spain(2007), Sweden(2002), Switzterland(2002, 2007).

6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}
library(plyr)
gapminder %>%
  data.frame(group = country, a = lifeExp , b = gdpPercap) %>%
require(plyr)
func <- function(xx)
return(data.frame(COR = cor(a, b)))
cor<-ddply(xx, .(group), func)
arrange(cor, desc(COR))

library(plyr)
xx <- data.frame(group = gapminder$country, a = gapminder$lifeExp , b = gapminder$gdpPercap)
head(xx)
require(plyr)
func <- function(xx)
return(data.frame(COR = cor(xx$a, xx$b)))
cor<-ddply(xx, .(group), func)
arrange(cor, desc(abs(COR))
```

7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).
```{r}
gapminder %>%
  group_by(continent, year) %>%
  filter(!continent == "Asia") %>%
  summarise(avg_pop = sum(pop)) %>%
  ungroup() %>%
  select(continent, country, pop, year) %>%
  arrange(pop)
```
ANSWER: Did not not solve :(. 

8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 
```{r}
gapminder %>%
  group_by(country) %>%
  select(country, pop) %>%
  aggregate(by = list(country), FUN = sum)
  summarise(total_pop = sum(pop))
  mutate(sd = sd(total_pop)) %>%
  select(country, sd)
```
ANSWER: Didn't solve :(. 

9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
```{r}
gm1992 <- gapminder %>%
  filter(year == "1992")
typeof(gm1992)
```
ANSWER: It is a list.

10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
```{r}

```
