# Grace pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
f3 <- function(x, y) {
  rep(y, length.out = length(x))
}
#Testing f1
df <- c("lets", "go")
d <- "le"
f1(df, d)
#Testing f2
p <- c(1,2,4, 3)
q <- 2
f2(q)
f2(p)
#Testing f3
f3(c(1,2,3,5,6,8), c(20, 19))
f3(c(19 ,19), c(1,2,3,5,6,8))
```
f1 rename prefix. It returns TRUE or FALSE if the prefix is in the string. 
f2 rename removelastobs. If vector has length of one or less, return NULL. Else, remove last observation in the vector. 
f3 rename samelength. It makes the second argument have the same length as the first argument in the function.
2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?
```{r}
?rnorm()
?MASS::mvrnorm()
```
rnorm samples from the univariate normal distribution. Its arguments are n, mean, sd.
MASS::mvrnorm samples from the multivariate normal distribution. Its arguments are n, mu, sigma.
To make them more consistent, they should have the same argument names. Sd does not, however, make sense for multivariate models. Therefore, I think it is ok for the different names because they accomplish the same thing from differing samples. 

3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 
```{r}
lapply(mtcars, function(x) sd(x)/mean(x))
```
4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    ```{r}
    vapply(x, sd)
    ```
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
    ```{r}
    vapply(vapply(x, as.numeric), sd)
    ```
